Giant Low Surface Brightness Galaxies As Merger Remnants

### 2015-05-05 glsb.tex
fill out intro

Observers tend to classify GLSB systems as normal spirals.  \citet{Matthews01b} list Malin~2 as an Scd while \citet{Das13} concludes that GLSB systems are an extreme form of late type spiral galaxies.  Similarly, \citet{Kasparova14} claim that the large dark matter scale of Malin~2 can explain it's features.  Even the Galaxy Zoo project classifies GLSB galaxies as spirals.  While GLSB systems superficially resemble normal spiral galaxies, they are clearly extreme outliers when their scale lengths are compared to other galaxy samples, being around a factor of 2 larger than the largest typical spirals \citep{Dalcanton97}.

The disks in GLSB systems are incredibly large, with Malin~1 being detected to a radius of over 100 kpc \citep{Moore06}.  Disks of this size are a challenge for typical $\Lambda$CDM cosmology galaxy formation and exotic formation scenarios have been proposed for GLSB systems.  \citet{Mapelli08} form GLSB-like galaxies via the fading of collisional ring galaxies.  In their model, a collision sets off an expanding ring of star formation.  As this the star formation subsides and the stellar population ages, the ring galaxy fades over $\sim$100 Myr to resembles observed GLSBs.  XXX-one problem is that most GLSB's are relatively isolated, and thus there are no obvious candidates for the impactor systems that would have triggered the star formation burst.

XXX-add discussion of possibility the centers of GLSBs are normal galaxies.

\citet{Barth07} analyze HST imaging of Malin 1 and find that the central object is well described as a normal SB0/a galaxy.  Makes good point that lots of the other large disk observations look like they are extensions of the inner galaxy, while Malin 1's disk is probably a distinct component.

>>>>

Observers tend to classify GLSB systems as normal spirals.  \citet{Matthews01b} list Malin~2 as an Scd while \citet{Das13} concludes that GLSB systems are an extreme form of late type spiral galaxies.  Similarly, \citet{Kasparova14} claim that the large dark matter scale of Malin~2 can explain it's features.  Even the Galaxy Zoo project classifies GLSB galaxies as spirals.  While GLSB systems superficially resemble normal spiral galaxies, they are clearly extreme outliers when their scale lengths are compared to other galaxy samples, being around a factor of 2 larger than the largest typical spirals \citep{Dalcanton97}. (xxx-maybe make a plot of SDSS spiral scale lengths and plot the GLSB systems).

The disks in GLSB systems are incredibly large, with Malin~1 being detected to a radius of over 100 kpc \citep{Moore06}.  Disks of this size are a challenge for typical $\Lambda$CDM cosmology galaxy formation and exotic formation scenarios have been proposed for GLSB systems.  \citet{Mapelli08} form GLSB-like galaxies via the fading of collisional ring galaxies.  In their model, a collision sets off an expanding ring of star formation.  As this the star formation subsides and the stellar population ages, the ring galaxy fades over $\sim$100 Myr to resembles observed GLSBs.  One obvious problem with this scenario is that most GLSB's are relatively isolated, and thus there are no obvious candidates for the impactor systems that would have triggered the star formation burst.

The final formation scenario for GLSB systems is that they are normal galaxies that undergo a high angular momentum minor merger where the LSB disk is the shredded remains of an accreted galaxy.  \citet{Pen06} find that extended disks can be formed via the disruption of dwarf galaxies.  These simulated debrie disks have exponential profiles and extend out to 200 kpc from the central galaxy.

Observations of Malin~1 support the accretion model of GLSBs.  \citet{Barth07} analyze HST imaging of Malin~1 and find that the central object is well described as a normal early-type barred galaxy that was previously unresolved in ground-based imaging. \citet{Barth07} points out that exteded disk structures in galaxies such as M31 appear to be a smooth continuation of the inner disk \citep{Ferguson02}, while Malin~1's extended disk is better described as a distinct component, (although \citet{Ibata05} claim M31's extended disk is of accretion origin).

If GLSB disks are indeed the remains of accreted dwarf galaxies, they represent a unique opportunity to observationally probe the outer regions of galaxy dark matter halos, especailly if they form around elliptical galaxies.  The mass profiles of elliptical galaxies have been difficult to constrain since they lack the HI disks found around spirals.
